---
title: "Activity Monitoring Analysis"
author: "Priyanka Dhamija"
date: "January 14, 2018"
output: html_document
---

#Read the data
```{r,echo = TRUE, results= TRUE, warning=FALSE}
rm(list=ls())
am<- read.csv(file.choose(), header = TRUE)
head(am,5)
```

#Total Missing values 
```{r, echo = TRUE, results= TRUE, warning=FALSE }
library(mice)
md.pattern(am) #2304 missing values in steps
colSums(is.na(am))
```

#Summary and structure of data set
```{r, echo = TRUE, results= TRUE, warning=FALSE }
summary(am)
str(am) #Date is in factor format
am$date<-as.Date(am$date)
```

#Calculate total number of steps each day
```{r, echo = TRUE, results= TRUE, warning=FALSE }
dailystepsmissing<-as.data.frame.table(tapply(am$steps, format(am$date, '%Y-%m-%d'), sum))
str(dailystepsmissing) #Convert date to date class
dailystepsmissing$Var1<-as.Date(dailystepsmissing$Var1)
```

#Histogram of total number of steps each day
```{r, echo = TRUE , results= TRUE, warning=FALSE}
library(ggplot2)
ggplot(dailystepsmissing, aes(x=Var1, y=Freq))+geom_histogram(stat="identity", fill="salmon")
```

# Average/Mean number of steps taken each day
```{r, echo = TRUE, results= TRUE, warning=FALSE }
library(dplyr)
mean_steps<- as.data.frame( am %>% group_by(date) %>% summarise(mean(steps)))
head(mean_steps,10)
```

#Median steps taken each day
```{r, echo = TRUE , results= TRUE, warning=FALSE}
median_steps<- as.data.frame( am %>% group_by(date) %>% summarise(median(steps,na.rm = FALSE)))
head(median_steps,10)
```

#Time Series plot of total number of steps without imputing Missing Values 
```{r, echo = TRUE , results= TRUE, warning=FALSE}
library(plotly)
p1 <- plot_ly(x = dailystepsmissing$Var1, y = dailystepsmissing$Freq, mode = 'lines') %>%
  layout(
    title = "Histogram of daily Steps with Missing Values"
  ) 
show(p1)
```

#5 minute interval that contains maximum number of steps each day
#Use aggregate to create new data frame with the maxima
```{r, echo = TRUE , results= TRUE, warning=FALSE}
am_agg <- aggregate(steps ~ date, am, max)
# then simply merge with the original
am_max <- merge(am_agg, am)
head(am_max,10)
```

#Using mice package to impute missings
#Mice package is a machine learning package which will consider other numeric features to impute 
#the missing package
```{r, echo = TRUE, results= TRUE, warning=FALSE }
am1<-am[,c(1,3)]
set.seed(123)
index1<- sample(2, nrow(am1), replace= T, prob = c(0.7,0.3))
library(mice)
init <- mice(am1, maxit=0) 
meth <- init$method
predM <- init$predictorMatrix
set.seed(100)
imputed = mice(am1, method=meth, predictorMatrix=predM, m=5)
imputed <- complete(imputed)
sapply(imputed, function(x) sum(is.na(x)))
am_new<-am
am_new$steps<-imputed$steps
summary(am_new)
```

#Calculate total number of steps each day
```{r, echo=TRUE,results= TRUE, warning=FALSE}
dailysteps<-as.data.frame.table(tapply(am$steps, format(am$date, '%Y-%m-%d'), sum))
str(dailysteps) #Convert date to date class
dailysteps$Var1<-as.Date(dailysteps$Var1)
```

#Histogram using ggplot
```{r, echo = TRUE, results= TRUE, warning=FALSE }
library(ggplot2)
ggplot(dailysteps, aes(x = dailysteps$Var1, y = dailysteps$Freq)) +
  geom_bar(stat = "identity", fill = "salmon")+ xlab('Date')+ylab('Number of steps')
```

#Panel Plot containing average number of steps taken per 5-minute interval across weekdays and weekends

#Create a function to separate weekedays and weekends
```{r, echo = TRUE, results= TRUE, warning=FALSE }
separation <- function(date) {
  if (weekdays(as.Date(date)) %in% c("Saturday", "Sunday")) {
    "weekend"
  } else {
    "weekday"
  }
}
am_new$wday <- as.factor(sapply(am_new$date, FUN=separation))
str(am_new)
```


#Panel Plot containing average number of steps taken per 5-minute interval across weekdays and weekends
```{r, echo = TRUE, results= TRUE, warning=FALSE }
am_new1 <- aggregate(am_new$steps, by=list(am_new$interval, am_new$wday),mean)
names(am_new1) <- c("interval", "wday", "steps")
xyplot(steps ~ interval | wday, am_new1, type = "l", layout = c(2,1),
       xlab = "Interval", ylab = "Average Number of Steps")
```




